---
title: "Binomial-Quabbin-Stats"
author: "Abby Robinson"
date: "1/11/2022"
output: html_document
---

```{r}
library(curl)
```

Now lets load the **Count Data** into our markdown file using the {curl} command 

```{r}
c <- curl("https://raw.githubusercontent.com/butterfliesrcool/Quabbin_Data/main/learning-phase-transect-one.csv")
l1  <- read.csv(c, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(l1) #we can use the head function to view the first 10 rows in the dataset 
length(l1$attacks)
sum(l1$attacks)
```

```{r}
t1 <- rep(1, each=4000)
l1$transect <- t1
head(l1)
```

```{r}
r <- curl("https://raw.githubusercontent.com/butterfliesrcool/Quabbin_Data/main/learning-phase-transect-two.csv")
l2  <- read.csv(r, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(l2) #we can use the head function to view the first 10 rows in the dataset 
length(l2$attacks)
sum(l2$attacks)
```

```{r}
t2 <- rep(2, each=4000)
l2$transect <- t2
head(l2)
```

```{r}
p <- curl("https://raw.githubusercontent.com/butterfliesrcool/Quabbin_Data/main/learning-phase-transect-three.csv")
l3  <- read.csv(p, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(l3) #we can use the head function to view the first 10 rows in the dataset 
length(l3$attacks)
sum(l3$attacks)
```

```{r}
t3 <- rep(3, each=4000)
l3$transect <- t3
head(l3)
```

```{r}
learning <- rbind(l1, l2, l3)
length(learning$attacks)
sum(learning$attacks)
```

```{r}
no.zero.learning <- subset(learning, attacks == "1", drop = FALSE)
table(no.zero.learning$attacks, no.zero.learning$species)
chisq.test(table(no.zero.learning$attacks, no.zero.learning$species))
```

```{r}
lb <- subset(no.zero.learning, species == "battus", drop = FALSE)
table(lb$attacks, lb$day)
chisq.test(table(lb$attacks, lb$day))
```

```{r}
lj <- subset(no.zero.learning, species == "junonia.learning", drop = FALSE)
table(lj$attacks, lj$day)
chisq.test(table(lj$attacks, lj$day))
```

```{r}
pairs(learning)
glm <- glm(data = learning, attacks ~ species + day, family = "binomial")
summary(glm)
confint(glm)
```

```{r}
b <- subset(learning, species == "battus", drop = FALSE)
glm <- glm(data = b, attacks ~ day, family = "poisson")
summary(glm)
confint(glm)
```

```{r}
j <- subset(learning, species == "junonia.learning", drop = FALSE)
glm <- glm(data = j, attacks ~ day, family = "poisson")
summary(glm)
confint(glm)
```

##run GLMs comparing species only for days 2-4 (this would work, theoretically, if birds learn in one day)

```{r}
##pld = post learning days 
pld <- subset(learning, day %in% c(2,3,4), drop = FALSE)
pld
table(pld$attacks, pld$species)

glm <- glm(data = pld, attacks ~ species, family = "poisson")
summary(glm)
```
##totally attacks excluding day one (where no learning has occurred) is significantly different between junonia and battus according to the GLM analysis 


```{r}
library(survival)
library(survminer)
library(ggplot2)
```

```{r}
model <- survfit(Surv(day, day+1, attacks) ~ species, data=learning)
model
summary(model)
cox.zph(model)

model2 <- survfit(Surv(day, day+1, attacks) ~ species, data=no.zero.learning)
coxph(Surv(day, day+1, attacks) ~ species, data=no.zero.learning)
survdiff(Surv(day, attacks) ~ day, data=no.zero.learning)
```

```{r}
p <- ggsurvplot(fit = model,
           xaxt = "n",
           xlab="Time (days)", 
           ylab="Survivorship",
           ylim = c(0.97, 1.0), 
           xlim = c(1, 5),  
           data = learning, 
           axes = FALSE, 
           linetype = 1,
           legend.title = "Species")

p$plot <- p$plot + theme_minimal()
p$plot <- p$plot + theme_classic()

##rename the x axis values 
p$plot <- p$plot + scale_x_continuous(breaks = c(1,2,3,4,5), labels = c(0,1,2,3,4))

p$plot <- p$plot + scale_color_manual(values = c("darkblue", "sienna4"))

p$plot
```



##Testing Phase Data input and analysis 
```{r}
a <- curl("https://raw.githubusercontent.com/butterfliesrcool/Quabbin_Data/main/testing-phase-transect-one.csv")
te1  <- read.csv(a, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(te1) #we can use the head function to view the first 10 rows in the dataset 
length(te1$attacks)
sum(te1$attacks)
```

```{r}
tt1 <- rep(1, each=4000)
te1$transect <- tt1
head(te1)
```

```{r}
h <- curl("https://raw.githubusercontent.com/butterfliesrcool/Quabbin_Data/main/testing-phase-transect-two.csv")
te2  <- read.csv(h, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(te2) #we can use the head function to view the first 10 rows in the dataset 
length(te2$attacks)
sum(te2$attacks)
```

```{r}
tt2 <- rep(2, each=4000)
te2$transect <- tt2
head(te2)
```

```{r}
w <- curl("https://raw.githubusercontent.com/butterfliesrcool/Quabbin_Data/main/testing-phase-transect-three.csv")
te3  <- read.csv(w, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(te3) #we can use the head function to view the first 10 rows in the dataset 
length(te3$attacks)
sum(te3$attacks)
```

```{r}
tt3 <- rep(3, each=4000)
te3$transect <- tt3
head(te3)
```
```{r}
testing <- rbind(te1, te2, te3)
length(testing$attacks)
sum(testing$attacks)
```


```{r}
no.zero.testing <- subset(testing, attacks == "1", drop = FALSE)
table(no.zero.testing$attacks, no.zero.testing$species)
chisq.test(table(no.zero.testing$attacks, no.zero.testing$species))
```

```{r}
tl <- subset(no.zero.testing, species == "limenitis", drop = FALSE)
table(tl$attacks, tl$day)
chisq.test(table(tl$attacks, tl$day))
```

```{r}
tj <- subset(no.zero.testing, species == "junonia", drop = FALSE)
table(tj$attacks, tj$day)
chisq.test(table(tj$attacks, tj$day))
```

```{r}
glm2 <- glm(data = testing, attacks ~ species + day, family = "poisson")
summary(glm2)
confint(glm2)
```

```{r}
l <- subset(testing, species == "limenitis", drop = FALSE)
glm3 <- glm(data = l, attacks ~ day, family = "poisson")
summary(glm3)
confint(glm3)
```

```{r}
tj <- subset(testing, species == "junonia", drop = FALSE)
glm4 <- glm(data = tj, attacks ~ day, family = "poisson")
summary(glm4)
confint(glm4)
```

##run GLMs comparing species only for days 2-4 (this would work, theoretically, if birds learn in one day)

```{r}
##pld = post learning days 
pld.testing <- subset(testing, day %in% c(2,3,4), drop = FALSE)
pld.testing
table(pld.testing$attacks, pld.testing$species)

glm <- glm(data = pld.testing, attacks ~ species, family = "poisson")
summary(glm)
##attacks excluding day one (where no learning has occurred) are not significantly different between junonia and limenitis 
```

```{r}
model2 <- survfit(Surv(day, day+1, attacks) ~ species, data=testing)
model2
summary(model2)
```

```{r}
p <- ggsurvplot(fit = model2,
           xaxt = "n",
           xlab="Time (days)", 
           ylab="Survivorship",
           ylim = c(0.98, 1.0), 
           xlim = c(1, 5),  
           data = testing, 
           axes = FALSE, 
           linetype = 1,
           legend.title = "Species")

p$plot <- p$plot + theme_minimal()
p$plot <- p$plot + theme_classic()

##rename the x axis values 
p$plot <- p$plot + scale_x_continuous(breaks = c(1,2,3,4,5), labels = c(0,1,2,3,4))

p$plot <- p$plot + scale_color_manual(values = c("sienna4", "deepskyblue"))

p$plot
```


creating stations could reduce the sample size and help eliminate noise in the data - this didn't seem to change anything

```{r}
table(learning$attacks, learning$fac.no) 
stations <- subset(learning, fac.no %in% c(1, 2, 3), drop = FALSE)
length(stations$attacks)
```

```{r}
bat.stations <- subset(stations, species == "battus", drop = FALSE)
```

```{r}
coxph(Surv(day, attacks) ~ 1, data=bat.stations)
survdiff(Surv(day, attacks) ~ day, data=bat.stations)
```

```{r}
jun.stations <- subset(stations, species == "junonia", drop = FALSE)
coxph(Surv(day, attacks) ~ 1, data=jun.stations)
survdiff(Surv(day, attacks) ~ day, data=jun.stations)
```

```{r}
coxph(Surv(day, day+1, attacks) ~ species, data=stations)
```

```{r}
survdiff(Surv(day, attacks) ~ species, data=stations)
pairwise_survdiff(Surv(day, attacks) ~ species, p.adjust.method = "BH",
     data = stations)
```


##Simultaneous data 

```{r}
u <- curl("https://raw.githubusercontent.com/butterfliesrcool/Quabbin_Data/main/simultaneous-transect-four.csv")
sim  <- read.csv(u, header = TRUE, sep = ",", stringsAsFactors = FALSE)
```
```{r}
no.zero.sim <- subset(sim, attacks == "1", drop = FALSE)
table(no.zero.sim$attacks, no.zero.sim$species)
glm <- glm(data = sim, attacks ~ species, family = "poisson")
summary(glm)
```

###GLMs for simultaneous experiment not significant - probably because samle size is too low 
NOTE FOR NEXT TIME: if we are going to treat the simultaneous experiment as a completely separate experiment, we should do one paired with each transect so that the sample sizes are comparable 

```{r}
##pld = post learning days 
pld.sim <- subset(sim, day %in% c(2,3,4), drop = FALSE)
table(pld.sim$attacks, pld.sim$species)

glm <- glm(data = pld.sim, attacks ~ species, family = "poisson")
summary(glm)
```

Attempts at coxme analysis 

```{r}
library(survival)
library(coxme)
###how do i test if the data meets the assumptions of the model??? 
### run the code "cox.zph()" to test whether the proportional hazards assumption is met 
all.data <- rbind(learning, testing)
length(all.data$attacks)
head(all.data)
```

```{r}
stem(table(eortc$center))
```

```{r}
eortc
```


```{r}
efit1 <- coxph(Surv(y, uncens) ~ trt, eortc)
efit2 <- coxme(Surv(y, uncens) ~ trt + (1|center), eortc)
```

```{r}
print(efit2)
```

```{r}
m1 <- coxph(Surv(day, day+1, attacks) ~ species + cluster(site + fac.no), 
    data = all.data)
summary(m1)
cox.zph(m1)
```

```{r}
m2 <- coxme(Surv(day, attacks) ~ species + day + (1|site), data = all.data)
summary(m2)
```

```{r}
m3 <- coxme(Surv(day, attacks) ~ species + day + (1|fac.no), data = all.data)
summary(m3)
```

```{r}
m4 <- coxme(Surv(day, attacks) ~ species + day + (1|site/fac.no), data = all.data)
summary(m4)
```


