---
title: "final.quabbin.analysis"
author: "Abby Robinson"
date: "1/11/2022"
output: html_document
---

```{r}
library(curl)
```


```{r}
c <- curl("https://raw.githubusercontent.com/butterfliesrcool/Quabbin_Data/main/learning-phase-transect-one.csv")
l1  <- read.csv(c, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(l1) #we can use the head function to view the first 10 rows in the dataset 
length(l1$attacks)
sum(l1$attacks)
```

```{r}
t1 <- rep(1, each=4000)
l1$transect <- t1
head(l1)
```

```{r}
r <- curl("https://raw.githubusercontent.com/butterfliesrcool/Quabbin_Data/main/learning-phase-transect-two.csv")
l2  <- read.csv(r, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(l2) #we can use the head function to view the first 10 rows in the dataset 
length(l2$attacks)
sum(l2$attacks)
```

```{r}
t2 <- rep(2, each=4000)
l2$transect <- t2
head(l2)
```

```{r}
p <- curl("https://raw.githubusercontent.com/butterfliesrcool/Quabbin_Data/main/learning-phase-transect-three.csv")
l3  <- read.csv(p, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(l3) #we can use the head function to view the first 10 rows in the dataset 
length(l3$attacks)
sum(l3$attacks)
```

```{r}
t3 <- rep(3, each=4000)
l3$transect <- t3
head(l3)
```

```{r}
learning <- rbind(l1, l2, l3)
length(learning$attacks)
sum(learning$attacks)
```


```{r}
glm <- glm(data = learning, attacks ~ species, family = "binomial")
summary(glm)
confint(glm)
```

```{r}
b <- subset(learning, species == "battus", drop = FALSE)
glm <- glm(data = b, attacks ~ day, family = "binomial")
summary(glm)
confint(glm)
```

```{r}
j <- subset(learning, species == "junonia.learning", drop = FALSE)
glm <- glm(data = j, attacks ~ day, family = "binomial")
summary(glm)
confint(glm)
```

##run GLMs comparing species only for days 2-4 (we would expect to see lower attack rates only on days 2-4 because predators are naive on day 1)

```{r}
##pld = post learning days 
pld <- subset(learning, day %in% c(2,3,4), drop = FALSE)
pld
table(pld$attacks, pld$species)

glm <- glm(data = pld, attacks ~ species, family = "binomial")
summary(glm)
```
##totally attacks excluding day one (where no learning has occurred) is significantly different between junonia and battus according to the GLM analysis 

##Testing Phase Data input and analysis 
```{r}
a <- curl("https://raw.githubusercontent.com/butterfliesrcool/Quabbin_Data/main/testing-phase-transect-one.csv")
te1  <- read.csv(a, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(te1) #we can use the head function to view the first 10 rows in the dataset 
length(te1$attacks)
sum(te1$attacks)
```

```{r}
tt1 <- rep(1, each=4000)
te1$transect <- tt1
head(te1)
```

```{r}
h <- curl("https://raw.githubusercontent.com/butterfliesrcool/Quabbin_Data/main/testing-phase-transect-two.csv")
te2  <- read.csv(h, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(te2) #we can use the head function to view the first 10 rows in the dataset 
length(te2$attacks)
sum(te2$attacks)
```

```{r}
tt2 <- rep(2, each=4000)
te2$transect <- tt2
head(te2)
```

```{r}
w <- curl("https://raw.githubusercontent.com/butterfliesrcool/Quabbin_Data/main/testing-phase-transect-three.csv")
te3  <- read.csv(w, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(te3) #we can use the head function to view the first 10 rows in the dataset 
length(te3$attacks)
sum(te3$attacks)
```

```{r}
tt3 <- rep(3, each=4000)
te3$transect <- tt3
head(te3)
```

```{r}
testing <- rbind(te1, te2, te3)
length(testing$attacks)
sum(testing$attacks)
```

```{r}
glm2 <- glm(data = testing, attacks ~ species, family = "binomial")
summary(glm2)
confint(glm2)
```

```{r}
l <- subset(testing, species == "limenitis", drop = FALSE)
glm3 <- glm(data = l, attacks ~ day, family = "binomial")
summary(glm3)
confint(glm3)
```

```{r}
tj <- subset(testing, species == "junonia", drop = FALSE)
glm4 <- glm(data = tj, attacks ~ day, family = "binomial")
summary(glm4)
confint(glm4)
```

##run GLMs comparing species only for days 2-4 


```{r}
##pld = post learning days 
pld.testing <- subset(testing, day %in% c(2,3,4), drop = FALSE)
pld.testing
table(pld.testing$attacks, pld.testing$species)

glm <- glm(data = pld.testing, attacks ~ species, family = "binomial")
summary(glm)
##attacks excluding day one (where no learning has occurred) are not significantly different between junonia and limenitis 
```


```{r}
p <- ggsurvplot(fit = model2,
           xaxt = "n",
           xlab="Time (days)", 
           ylab="Survivorship",
           ylim = c(0.98, 1.0), 
           xlim = c(1, 5),  
           data = testing, 
           axes = FALSE, 
           linetype = 1,
           legend.title = "Species")

p$plot <- p$plot + theme_minimal()
p$plot <- p$plot + theme_classic()

##rename the x axis values 
p$plot <- p$plot + scale_x_continuous(breaks = c(1,2,3,4,5), labels = c(0,1,2,3,4))

p$plot <- p$plot + scale_color_manual(values = c("sienna4", "deepskyblue"))

p$plot
```


creating stations could reduce the sample size and help eliminate noise in the data - this didn't seem to change anything

```{r}
table(learning$attacks, learning$fac.no) 
stations <- subset(learning, fac.no %in% c(1, 2, 3), drop = FALSE)
length(stations$attacks)
```

```{r}
bat.stations <- subset(stations, species == "battus", drop = FALSE)
```

```{r}
coxph(Surv(day, attacks) ~ 1, data=bat.stations)
survdiff(Surv(day, attacks) ~ day, data=bat.stations)
```

```{r}
jun.stations <- subset(stations, species == "junonia", drop = FALSE)
coxph(Surv(day, attacks) ~ 1, data=jun.stations)
survdiff(Surv(day, attacks) ~ day, data=jun.stations)
```

```{r}
coxph(Surv(day, day+1, attacks) ~ species, data=stations)
```

```{r}
survdiff(Surv(day, attacks) ~ species, data=stations)
pairwise_survdiff(Surv(day, attacks) ~ species, p.adjust.method = "BH",
     data = stations)
```


##Simultaneous data 

```{r}
u <- curl("https://raw.githubusercontent.com/butterfliesrcool/Quabbin_Data/main/simultaneous-transect-four.csv")
sim  <- read.csv(u, header = TRUE, sep = ",", stringsAsFactors = FALSE)
```


###GLMs for simultaneous experiment not significant - probably because samle size is too low 

##run GLMs comparing species only for days 2-4 

##totally attacks excluding day one (where no learning has occurred) is significantly different between junonia and battus according to the GLM analysis 
```{r}
##pld = post learning days 
pld.testing <- subset(testing, day %in% c(2,3,4), drop = FALSE)
pld.testing
table(pld.testing$attacks, pld.testing$species)

glm <- glm(data = pld.testing, attacks ~ species, family = "poisson")
summary(glm)
##attacks excluding day one (where no learning has occurred) are not significantly different between junonia and limenitis 
```

```{r}
attacks <- subset(sim, attacks == "1", drop = FALSE)
ba <- subset(sim, species == "battus", drop = FALSE)
li <- subset(sim, species == "limenitis", drop = FALSE)
ju <- subset(sim, species == "junonia", drop = FALSE)

bl <- subset(sim, species %in% c("battus", "limenitis"), drop = FALSE)
jl <- subset(sim, species %in% c("junonia", "limenitis"), drop = FALSE)
bj <- subset(sim, species %in% c("battus", "junonia"), drop = FALSE)
```

```{r}

glm4 <- glm(data = sim, attacks ~ species, family = "binomial")
summary(glm4)

glm.bl <- glm(data = bl, attacks ~ species, family = "binomial")
summary(glm.bl)

glm.jl <- glm(data = jl, attacks ~ species, family = "binomial")
summary(glm.jl)

glm.bj <- glm(data = bj, attacks ~ species, family = "binomial")
summary(glm.bj)
```

```{r}
table(attacks$species, attacks$attacks)
sim <- chisq.test(table(attacks$species, attacks$attacks))

x <- c("Battus", "Limenitis", "Junonia")
y <- c(7,4,15)
plot(x,y)
```


Attempts at coxme analysis 

```{r}
library(survival)
library(coxme)
###how do i test if the data meets the assumptions of the model??? 
### run the code "cox.zph()" to test whether the proportional hazards assumption is met 
all.data <- rbind(learning, testing)
length(all.data$attacks)
head(all.data)
```

```{r}
stem(table(eortc$center))
```

```{r}
eortc
```


```{r}
efit1 <- coxph(Surv(y, uncens) ~ trt, eortc)
efit2 <- coxme(Surv(y, uncens) ~ trt + (1|center), eortc)
```

```{r}
print(efit2)
```

```{r}
m1 <- coxph(Surv(day, day+1, attacks) ~ species + cluster(site + fac.no), 
    data = all.data)
summary(m1)
cox.zph(m1)
```

```{r}
m2 <- coxme(Surv(day, attacks) ~ species + day + (1|site), data = all.data)
summary(m2)
```

```{r}
m3 <- coxme(Surv(day, attacks) ~ species + day + (1|fac.no), data = all.data)
summary(m3)
```

```{r}
m4 <- coxme(Surv(day, attacks) ~ species + day + (1|site/fac.no), data = all.data)
summary(m4)
```

###GLMs for simultaneous experiment not significant - probably because samle size is too low 
NOTE FOR NEXT TIME: if we are going to treat the simultaneous experiment as a completely separate experiment, we should do one paired with each transect so that the sample sizes are comparable 

##run GLMs comparing species only for days 2-4 (this would work, theoretically, if birds learn in one day)

```{r}
##learing phase day-to-day pairwise comparisons 
b1.2 <- subset(learning, day %in% c(1,2), drop = FALSE)
b1.3 <- subset(learning, day %in% c(1,3), drop = FALSE)
b1.4 <- subset(learning, day %in% c(1,4), drop = FALSE)
b2.3 <- subset(learning, day %in% c(2,3), drop = FALSE)
b2.4 <- subset(learning, day %in% c(2,4), drop = FALSE)
b3.4 <- subset(learning, day %in% c(3,4), drop = FALSE)
```

```{r}
glm.b1.2 <- glm(data = b1.2, attacks ~ species, family = "binomial")
summary(glm.b1.2)
##no significant difference in attack rates between species (p=0.66686)
##significant differnce in attack rates between day 1&2 (p=0.00157 ** )
```

```{r}
glm.b1.3 <- glm(data = b1.3, attacks ~ species, family = "binomial")
summary(glm.b1.3)
##no significant difference in attack rates between species (p=0.646)
##significant differnce in attack rates between day 1&2 (p=6.86e-05 ***)
```

```{r}
glm.b1.4 <- glm(data = b1.4, attacks ~ species + day, family = "binomial")
summary(glm.b1.4)
##no significant difference in attack rates between species (p=0.76700 )
##significant differnce in attack rates between day 1&2 (p=0.00032 ***)
```

```{r}
glm.b2.3 <- glm(data = b2.3, attacks ~ species, family = "binomial")
summary(glm.b2.3 )
##no significant difference in attack rates between species (p=0.0819)
##no significant differnce in attack rates between day 1&2 (p=0.1858)
```

##totally attacks excluding day one (where no learning has occurred) is significantly different between junonia and battus according to the GLM analysis 
```{r}
##pld = post learning days 
pld.testing <- subset(testing, day %in% c(2,3,4), drop = FALSE)
pld.testing
table(pld.testing$attacks, pld.testing$species)

glm <- glm(data = pld.testing, attacks ~ species, family = "poisson")
summary(glm)
##attacks excluding day one (where no learning has occurred) are not significantly different between junonia and limenitis 
```

x: a vector of counts of successes
n: a vector of count trials
alternative: a character string specifying the alternative hypothesis
correct: a logical indicating whether Yates’ continuity correction should be applied where possible

```{r}
all.attacks <- prop.test(x = c(28, 28), n = c(66, 48))
all.attacks
```

```{r}
post.learning <- prop.test(x = c(8, 16), n = c(30, 26))
post.learning
```

```{r}
T1 <- prop.test(x = c(11, 7), n = c(18, 18))
# Printing the results
T1
```

```{r}
T2 <- prop.test(x = c(10, 6), n = c(16, 16))
# Printing the results
T2
```

```{r}
T3 <- prop.test(x = c(7, 7), n = c(14, 14))
# Printing the results
T3
```

```{r}
sim <- prop.test(x = c(4, 15), n = c(19, 19))
# Printing the results
sim
```



```{r}
result <- c(T1$p.value, T2$p.value, T3$p.value, sim$p.value)
result <- signif(result, digits = 3)
result
```

```{r}
df_p_val <- data.frame(
  group1 = "1",
  group2 = "2",
  label = result,
  y.position = 6
)
```
