---
title: "LEARNING:quabbin_data"
author: "Abby_Robinson"
date: "11/8/2021"
output: html_document
---
```{r}
setwd("/Users/AbbyRobinson/Desktop/Summer 2021 Data")
data <- read.csv("yes.data.csv")
data
```
subset only learning (model) phase data
```{r}
l <- subset(data, phase == "model", drop = FALSE)
l
```

Attack totals by day for battus philenor 
```{r}
b <- subset(l, species == "battus", drop = FALSE)
b
aggregate(x= b$attacks, by= list(b$day), FUN=sum)

battus.aov <- aov(b$attacks ~ b$day)
summary(battus.aov)

chisq.test(aggregate(x= b$attacks, by= list(b$day), FUN=sum))
##WARNING: Chi-squared approximation may be incorrect
```

Attack totals by day for junonia coenia 
```{r}
j <- subset(l, species == "junonia", drop = FALSE)
j
aggregate(x= j$attacks, by= list(j$day), FUN=sum)

junonia.aov <- aov(j$attacks ~ j$day)
summary(junonia.aov)

chisq.test(aggregate(x= j$attacks, by= list(j$day), FUN=sum)) 
##WARNING: Chi-squared approximation may be incorrect
```

Graphed attack rates for all learning phase transects (summed) based on raw totals
```{r}
setwd("/Users/AbbyRobinson/Desktop/Summer 2021 Data")
attacks <- read.csv("survival.csv")
attacks

#Calculated survival rates for battus and junonia 

####battus 
(1500 - 20)/1500 #0.9866667 survival rate ##DAY 1
(1500 - 23)/1500 #0.9846667 survival rate ##DAY 2
(1500 - 26)/1500 #0.9826667 survival rate ##DAY 3
(1500 - 28)/1500 #0.9813333 survival rate ##DAY 4

####junonia 
(1500 - 16)/1500 #0.9893333 survival rate ##DAY 1
(1500 - 25)/1500 #0.9833333 survival rate ##DAY 2
(1500 - 29)/1500 #0.9806667 survival rate ##DAY 3
(1500 - 37)/1500 #0.9753333 survival rate ##DAY 4

library(ggplot2)

g <- ggplot(data=attacks, aes(x=day, y=survival, color=species))
g <- g + geom_step()
g <- g + geom_point()
g

##website: http://www.sthda.com/english/wiki/ggplot2-rotate-a-graph-reverse-and-flip-the-plot
```

Susan used the survival package to calculate and graph survival "scores" over time. 
**also do a Cox proportional-hazards regression model 
```{r}
setwd("/Users/AbbyRobinson/Desktop/Summer 2021 Data")
data <- read.csv("yes.data.csv")
data

library(survival)

l <- subset(data, phase == "model", drop = FALSE)
l

surv <- survfit(Surv(day, attacks) ~ species, data = l)
surv

#compare survival based on attack numbers over 4-days, "~ species" allows me to compare between species 
#Invalid status value, converted to NA

survdiff(Surv(day, attacks) ~ species, data= l) 
##This isn't working anymore... because its count data and not binomial data?
```

survival scores for each species graphed in ggplot
is it possible to extend the stair graph?
```{r}
library(ggplot2)
library(survival)
library(survminer)

g <- ggsurvplot(data = data, 
    fit = surv, 
    xlab = "Days", 
    ylab = "survival",
    ylim = c(0.75, 1.00))
g
```


