---
title: "log.likelihood.code"
author: "Abby Robinson"
date: "2/5/2022"
output: html_document
---

This technique estimates the probability (QA i × ju) that avian predators in each site would attack a facsimile of type i relative to that of the Junonia control facsimile ju, by setting the attack rate on ju as a theoretical maximum such that QA i = (XA i /(XA ju + XA i )). Thus, for predation between Junonia controls versus other species i facsimiles, the actual probabilities are QA i×ju/(QA i×ju + 1) that predators attack i and 1/(QA i×ju + 1) that they attack ju. 

Attacks on Battus during learning phase (transects 1-3 combined)
DAY 1: 20
DAY 2: 3
DAY 3: 3
DAY 4: 2

Attacks on Junonia during learning phase (transects 1-3 combined)
DAY 1: 16
DAY 2: 10
DAY 3: 4
DAY 4: 8

```{r}
##i = attacks on facsimile of a specific phenotype 
##jun = attacks on junonia 
log.likelihood = function(i, jun){
  
  prob <- ((i / (jun+i)) * (jun / (jun+i)))
  
  likelihood <- (i * log(prob / (prob+1)) ) + (jun * log(1/ (prob+1)) ) 
  
  cat("probability =", prob <- prob, "\n",
      "likelihood =", likelihood <- likelihood, "\n")
}
```

```{r}
##log.likelihood analysis for battus during learning phase for day 1
log.likelihood(20, 16)
bat.day.one <- -35.91851
```

```{r}
##log.likelihood analysis for battus during learning phase for day 2
log.likelihood(3, 10)
bat.day.two <- -7.310383 
```

```{r}
##log.likelihood analysis for battus during learning phase for day 3
log.likelihood(3, 4)
bat.day.three <- -5.754116 
```

```{r}
##log.likelihood analysis for battus during learning phase for day 3
log.likelihood(2, 8)
bat.day.four <- -5.149363 
```

GTest is then used to compare log likelihood values between days and between species and the resulting P values were adjusted by applying a Bonferroni cor- rection to guard against false positives from multiple comparison

```{r}
library(lme4)
library(DescTools) 

battus.likelihood <- rbind(c(abs(bat.day.one), abs(bat.day.two), abs(bat.day.three), abs(bat.day.four))) 
GTest(battus.likelihood, correct = "none")
```

