---
title: "TESTING_quabbin_data"
author: "Abby_Robinson"
date: "11/16/2021"
output: html_document
---

Notes: 
1. do i need to show that attacks are not significantly different across 4 days in the field in order for this to show learning? (or the absence of learning...)

```{r}
setwd("/Users/AbbyRobinson/Desktop/Summer 2021 Data")
data <- read.csv("yes.data.csv")
data
```

```{r}
t <- subset(data, phase == "mimic", drop = FALSE)
t
```

#Testing Phase Attacks 

transect 1 attack totals & ANOVA (not significant)
```{r}
t1 <- subset(t, transect == "1", drop = FALSE)
t1
aggregate(x= t1$attacks, by= list(t1$species), FUN=sum)

t1.aov <- aov(t1$attacks ~ t1$species)
summary(t1.aov)
###ANOVA: attacks on the mimic are not significantly different from attacks on the control 
```

plotting attack totals
```{r}
species <- c("junonia", "limenitis")
attacks <- c(7, 11)

barplot(attacks ~ species)
```

junonia attack rate & attacks by day
```{r}
jun <- subset(t1, species == "junonia", drop = FALSE)
aggregate(x= jun$attacks, by= list(jun$day), FUN=sum)
sum(jun$attacks) #7 attacks on junonia
#500 facsimilies in each transect (20 sites, 25 junonia facsimilies in each site)
7/500 #1.4% attack rate on junonia 
```

limenitis attack rate & attacks by day
```{r}
lim <- subset(t1, species == "limenitis", drop = FALSE)
aggregate(x= lim$attacks, by= list(lim$day), FUN=sum)
sum(lim$attacks) #11 attacks on limenitis 
#500 facsimilies in each transect (20 sites, 25 junonia facsimilies in each site)
11/500 #2.2% attack rate on limenitis 
```

transect 2 attack totals & ANOVA (not significant)
```{r}
t2 <- subset(t, transect == "2", drop = FALSE)
t2
aggregate(x= t2$attacks, by= list(t2$species), FUN=sum)

t2.aov <- aov(t2$attacks ~ t2$species)
summary(t2.aov)
###ANOVA: attacks on the mimic are not significantly different from attacks on the control 
```

plotting attack totals
```{r}
species <- c("junonia", "limenitis")
attacks <- c(6, 10)

barplot(attacks ~ species)
```

junonia attack rate & attacks by day
```{r}
jun <- subset(t2, species == "junonia", drop = FALSE)
aggregate(x= jun$attacks, by= list(jun$day), FUN=sum)
sum(jun$attacks) #6 attacks on junonia
#500 facsimilies in each transect (20 sites, 25 junonia facsimilies in each site)
6/500 #1.2% attack rate on junonia 
```

limenitis attack rate & attacks by day
```{r}
lim <- subset(t2, species == "limenitis", drop = FALSE)
aggregate(x= lim$attacks, by= list(lim$day), FUN=sum)
sum(lim$attacks) #10 attacks on limenitis 
#500 facsimilies in each transect (20 sites, 25 junonia facsimilies in each site)
10/500 #2.0% attack rate on limenitis 
```

transect 3 attack totals & ANOVA (not significant)
```{r}
t3 <- subset(t, transect == "3", drop = FALSE)
t3
aggregate(x= t3$attacks, by= list(t3$species), FUN=sum)

t3.aov <- aov(t3$attacks ~ t3$species)
summary(t3.aov)
###ANOVA: attacks on the mimic are not significantly different from attacks on the control 
```

plotting attack totals
```{r}
species <- c("junonia", "limenitis")
attacks <- c(7, 7)

barplot(attacks ~ species)
```

junonia attack rate & attacks by day
```{r}
jun <- subset(t3, species == "junonia", drop = FALSE)
aggregate(x= jun$attacks, by= list(jun$day), FUN=sum)
sum(jun$attacks) #7 attacks on junonia
#500 facsimilies in each transect (20 sites, 25 junonia facsimilies in each site)
7/500 #1.4% attack rate on junonia 
```

limenitis attack rate & attacks by day
```{r}
lim <- subset(t3, species == "limenitis", drop = FALSE)
aggregate(x= lim$attacks, by= list(lim$day), FUN=sum)
sum(lim$attacks) #7 attacks on limenitis 
#500 facsimilies in each transect (20 sites, 25 junonia facsimilies in each site)
7/500 #1.4% attack rate on limenitis 
```


#Simultaneous experiment (transect 4)

```{r}
s <- subset(data, phase == "simultaneous", drop = FALSE)
s
```

```{r}
aggregate(x= s$attacks, by= list(s$species), FUN=sum)

sim.aov <- aov(s$attacks ~ s$species)
summary(sim.aov)
###Ok, so this isn't significant.... boo.... 
```

attack rates for simultaneous experiment 
```{r}
7/500 #battus 1.4% attack rate 

15/500 #junonia 3.0% attack rate

4/500 #limenitis 0.8% attack rate 
```

plotting attack totals
```{r}
species <- c("junonia", "battus", "limenitis")
attacks <- c(15, 7, 4)

barplot(attacks ~ species)
```

chi.squared isn't working with the count data anymore... 
```{r}
table(s$attacks, s$species)
test4 <- chisq.test(table(tp4$ATTACKS, tp4$SPECIES))
test4
#p-value = 0.01312! significant, right?? 
#do I need to do a post hoc test to show that attacks on battus and limenitis are significantly lower than attacks on junonia?
```

 
